---
title: "Simulation methods"
author: "Gwo"
date: "2024-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulation
## sample()
### Permutations 
  * it works on any object that has a defined length().
  * for example, sample(1:6); sample(list("A",3,sum))

### Resampling
  * Resampling from any existing distribution gets us the bootstrap family of
estimators
  * the jackknife worked by removing one point from the sample and recalculating the statistic of interest. Here resample the same length with replacement

```{r}
bootstrap.resample <- function (object){
sample (object, length(object), replace=TRUE)}
replicate(5, bootstrap.resample (6:10))
```

### bootstrap test 
  * The 2-sample t-test checks for differences in means according to a known null
distribution

### Markov Chain
  * Markov dependence:
  $p(X_1, X_2, ..., X_n) = p(X_1)p(X_2|X_1)...p(X_n|X_{n-1},...,X_1) = p(X_1)\prod_{i=2}^np(X_i|X_{i-1})$
  * generating a Markov Chain
    - Set up conditional distribution
    - Draw the initial state of the chain
    - For every additional draw, use the previous draw to inform the new one
  * simple unbounded Markov Chain
    - "Unbiased Random Walk": Independent events atop a dependent structure.
  