---
title: "Statistical models"
author: "Gwo"
date: "2024-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TestCase:Birth weightdata
```{r}
library(tidyverse)
library(MASS)
data(birthwt)
summary(birthwt)

# make ir readable
colnames(birthwt) <- c("birthwt.below.2500","mother.age","mother.weight","race","mother.smokes","previous.prem.labor","hypertension","uterine.irr","physician.visits","birthwt.grams")

# make all the factors more description
birthwt$race <- factor(c("white","black","other")[birthwt$race])
birthwt$mother.smokes <- factor(c("No", "Yes")[birthwt$mother.smokes + 1])
birthwt$uterine.irr <- factor(c("No", "Yes")[birthwt$uterine.irr + 1])
birthwt$hypertension <- factor(c("No", "Yes")[birthwt$hypertension + 1])
```

## plot to observe
### bar plot
### scatterplot
### boxplot

## Basic statistical methods
### linear model
  * t-test, p-value
  * lm(x~y,data)

### GeneralizedLinear Models
  * The default value is a Gaussian model (a standard linear model)
  * glm -- family = distribution(link=function)

## random number generators
  * runif() and set.seed()

## notes
  * the 0-1 variable needs +1 to correspond
  * <<-: Used to change variables in the global environment within a function,"super assignment"