---
title: "data science and application"
author: "Gwo"
date: "2024-04-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data type and structure
## Data type
  All data is represented in binary format, by bits (TRUE/FALSE, YES/NO, 1/0).
  * Booleans:TRUE or FALSE
  * Integers
  * Characters;strings = sequences of characters
  * Floating point numbers
  * Missing or ill-defined values: NA, NaN, etc
  
### Related function
  * typeof()
  * is.foo():whether the argument is of type foo
  * as.foo():"cast" its argument to type foo
```{r}
is.na(7/0)
is.na(0/0)
# since 7/0 is Inf or -Inf while 0/0 is NaN(not a number).
```

### Pecularites of floating-point numbers
  finite precision $\Rightarrow$ arithmetic on doubles $\neq$ on $R$.
  * Rounding errors tend to accumulate in long calculations
  * Usually better to use all.equal() than exact comparison

## Data structure
### vectors
  * A vector is a sequence of values, all of the same type;
  * Recycling repeat elements in shorter vector when combined with longer
  * To compare whole vectors, best to use identical() or all.equal()
  * R quantile：x% (n-1) + 1.
  
### arrays
  * Can have 3, 4,…n dimensional arrays; dim is a length n vector
  
### matrices
  * rowMeans(), colMeans(): input is matrix, output is vector
  * summary(): vector-style summary of column
  * apply(), takes 3 arguments: the array or matrix, then 1 for rows and 2 for
    columns, then name of the function to apply to each
    
### lists
  * Sequence of values, not necessarily all of the same type
  * Accessing pieces of lists: use [ ] or [[ ]] same as $(drops names and structures)
  * Add to lists with c()
  * Chop off the end of a list by setting the length to something smaller

### dataframes
  * Dataframe = the classic data table, n rows for cases, p columns for variables
  * add rows or columns to an array or data-frame with rbind() and cbind()
  
## Something else
  * eigen() finds eigenvalues and eigenvectors of a matrix
  * Recursion lets us build complicated data structures out of the simpler ones

# Visualization:ggplot2
## ggplot2
  * a coherent system for describing and building graphs
  * ggplot likes data in the ‘long’ format: i.e., a column for every dimension, and a
row for every observation
  * install.package() and library()

## plotting with ggplot2
### load data  
  * use the function read_csv()
  * head() or view() can look up the data

### 7 parameters of ggplot2
  ggplot(data = <DATA>) +
    <GEOM_FUNCTION>(
      mapping = aes(<MAPPINGS>),
      stat = <STAT>,
      position = <POSITION>
) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>

### build a ggplot
  * use the ggplot() function and bind the plot to a specific data frame using the data argument
  * add geoms – graphical representation of the data in the plot (points, lines,bars).
    * geom_point() for scatter plots, dot plots
    * geom_bar() for bar charts 
    * geom_line() for trend lines, time-series
  * Building plots iteratively
  * Customizingplots
    * add x,y axis lables and main title of the figure
    * choose a theme to replace the standard default theme 
    * change the size color and shape of the data
    * ......
  * Faceting:split one plot into multiple plots based on data
  * Geometricobjects(geoms):A geom is the geometrical object that a plot uses to represent data
    * bar charts use bar geoms, line charts use line geoms, boxplots use boxplot geoms
    * Scatterplots use the point geom
  * position = "dodge" places overlapping objects directly beside one another
  * use the ggsave() function to save the plot to a file
  
```{r}
  library(tidyverse)
  # head(iris)
  ggplot(data = iris)+
    geom_point(aes(x=iris$Sepal.Length,y=iris$Petal.Length,color=factor(Species),shape=factor(iris$Species)))+
    labs(x="Sepal_Length",
         y="Petal_Length",
         title = "relation among the species")+
    theme_bw()+
    theme(legend.title=element_blank())
```

```{r}
  # faceting tech
  ggplot(data=iris)+
    geom_point(aes(x=iris$Sepal.Length,y=iris$Petal.Length,color=iris$Species))+
    facet_wrap(~iris$Species,scales = "free")
  # 'scales' cancel the limit of axis
```
### notes
  * Anything putted in the ggplot() function can be seen by any geom layers
  * To display multiple geoms in the same plot, add multiple geom functions
  * Many graphs, like scatterplots, plot the raw values of dataset. Other graphs, like bar charts, calculate new values to plot
  * color a bar chart using either the color aesthetic, or, more usefully, fill.
  
# Tide data
## Data wrangling:dplyr
